---
title: "FACE Toxin & Fusarium load data"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, message=FALSE)
```

* Run using `r version[['version.string']] `.

# Objective
This document reports on the analysis of Toxin (ug DON/g tissue) and fungal biomass from Glenn wheat samples inoculated in the SoyFACE field experiment with GZ3639 in 2017 and 2018. 



## Details
Field experiment was conducted as part of SOY-FACE ambient and elevevated CO2 experiments at UIUC. Fungal biomass was constructed in 2017 from a single gene of Fusarium and single gene of wheat, using Bio_Rad, and expressed as a ratio. Fungal biomass in 2018 was constructed from multiple genes using Fluidigm, also expressed as a ratio.
DON concentrations were measured by Susan McCormick.


# 0) Load Packages, functions, source scripts
```{r, echo = FALSE, results = 'hide', include = FALSE} 
x <- c("car", "ggplot2", "dplyr", "lme4", "plyr", "reshape2", "tinytex", "egg")
lapply(x, require, character.only = TRUE)

#Load functions
source("./code/multiplot.function.R")
#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
#runif(1, min = 0, max = 1000)
set.seed(711)

# set ggplot2 theme
theme_set(theme_bw(base_size = 14)) 
theme_update(panel.grid.major = element_line(0), panel.grid.minor = element_line(0))


require(RColorBrewer)
display.brewer.pal(4, "Set1")
brewer.pal(4, "Set1")
co2_color <- c("#377EB8", "#E41A1C") 
tissue_color <- c("#4DAF4A", "#984EA3") 
display.brewer.pal(6, "BuPu")
brewer.pal(6, "BuPu")
collect_color <- c("#9EBCDA", "#8C96C6", "#8856A7", "#810F7C") 
```


# 1) Load data
```{r, results = 'hide'} 
yr2017 <- read.csv("./data/2017 FACE field data Biomass and DON.csv", stringsAsFactors = TRUE)
yr2018 <- read.csv("./data/2018 Data Biomass and Toxin.csv", stringsAsFactors = TRUE)

# convert treatments to factors
# NOTE - for now converting day collected in 2018 to a factor, not numeric
yr2017$year <- as.factor(yr2017$year); yr2017$plot <- as.factor(yr2017$plot); 
yr2017$replicate <- as.factor(yr2017$replicate)
yr2018$year <- as.factor(yr2018$year); yr2018$plot <- as.factor(yr2018$plot); 
yr2018$day <- as.factor(yr2018$day); yr2018$replicate <- as.factor(yr2018$replicate)
# add a DON per unit Fg biomass for year 2017
yr2017$DONperFg <- yr2017$DON/yr2017$FgTa_TRI6toPR1

# Calculate mean values per plot (only relevant to 2018)
yr2018.melted <- melt(yr2018, id.vars = c("CO2", "plot", "day"), measure.vars = c("FgTa_Geomean", "DON"), na.rm = TRUE)
yr2018.summarized <- ddply(yr2018.melted, .(CO2, plot, day, variable), summarise, averaged = mean(value, na.rm = TRUE))
yr2018.means <- dcast(yr2018.summarized, CO2 + plot + day ~ variable)

```


## 2) Visualize

### 2a) DON
```{r, echo = FALSE, results = 'hide'}
#expression(CO[2])

# DON
don17 <- ggplot(yr2017, aes(y = DON, x = CO2, group = CO2, color=CO2)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.5) +
    scale_y_continuous("DON (ppm)") +
    scale_x_discrete(expression(CO[2])) +
    guides(colour = 'none') +
    scale_color_manual("", values = co2_color)
don18 <- ggplot(yr2018) +
    facet_wrap(~ CO2, nrow = 1) +
    geom_boxplot(aes(y = DON, x = day, color=CO2), outlier.shape = NA) +
    geom_jitter(aes(y = DON, x = day, color=CO2), width = 0.1, height = 0, size = 1.5, alpha = 0.5)+
    scale_y_continuous("DON (ppm)") + 
    scale_x_discrete("Days Post Infection") +
    guides(colour = 'none') +
    scale_color_manual("", values = co2_color)

don17
don18
```

### 2b) Fusarium Biomass
```{r, echo = FALSE, results = 'hide'}
# NOTE 2017 data normally distributed, but 2018 definitely not
# can log transform both and have the 2017 data still look reasonably normal
fg17 <- ggplot(yr2017, aes(y = FgTa_TRI6toPR1, x = CO2, group = CO2, color=CO2)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(width = 0.1, height = 0, size = 2, alpha = 0.5) +
    scale_y_continuous(expression(paste(italic("Fusarium")," load")), 
                       trans = "log", breaks = c(2, 5, 10, 15)) +
    scale_x_discrete(expression(CO[2])) +
    guides(colour = 'none') +
    scale_color_manual("", values = co2_color)
fg18 <- ggplot(yr2018) +
    facet_wrap(~ CO2, nrow = 1) +
    geom_boxplot(aes(y = FgTa_Geomean, x = day, color=CO2), outlier.shape = NA) +
    geom_jitter(aes(y = FgTa_Geomean, x = day, color=CO2), width = 0.1, height = 0, size = 1.5, alpha = 0.5) +
    scale_y_continuous(expression(paste(italic("Fusarium")," load")), 
                       trans = "log", breaks = c(.5, 2, 5, 10)) + 
    scale_x_discrete("Days Post Infection") +
    guides(colour = 'none') +
    scale_color_manual("", values = co2_color)

fg17
fg18
```

```{r, include = FALSE}
fg_all <- ggarrange(fg17, fg18, don17, don18, 
          ncol = 2, nrow = 2,
          widths = c(1,1.5), heights = c(1,1),
          labels = c("A)", "B)", "C)", "D)"), 
          label.args = list(gp=gpar(font=2, fontsize=16), 
                            x=unit(1,"line"), vjust=1.5) )  #, hjust=-.5, 
#tiff("./figures/DON & Fg by CO2.tiff", width=180, height=160, units="mm", res=600)
fg_all
#dev.off()
#85mm width or 180mm width


# #tiff("./figures/DON & Fg by CO2.tiff", width=7, height=7, units="in", res=600)
# multiplot(fg17, fg18, don17, don18, cols = 2)
# #dev.off()
# #png("./figures/fg17.png", width = 6, height = 9, units = "cm", res = 1200)
# fg17
# #dev.off()
# #png("./figures/fg18.png", width = 12, height = 9, units = "cm", res = 1200)
# fg18
# #dev.off()
# #png("./figures/don17.png", width = 6, height = 9, units = "cm", res = 1200)
# don17
# #dev.off()
# #png("./figures/don18.png", width = 12, height = 9, units = "cm", res = 1200)
# don18
# #dev.off()

```


## 3) Stats
```{r, echo = FALSE, results = 'hide'}

### 2017 design
table(yr2017$CO2, yr2017$plot)
# plot nested in CO2, though with 1 plot per treatment, plot is defunct in this year, 
#   so really replicate nested in CO2

### 2018 design
table(yr2018$plot, yr2018$day, yr2018$CO2)
# plot nested in CO2, repeated measures expt


### normality
hist(yr2017$DON)  #odd, but also few data points, may need to try boxcox for this one
hist(sqrt(yr2018$DON)) #sqrt better than raw
hist(yr2017$FgTa_TRI6toPR1)  #fine as is, log doesn't appear to hurt, could be cautious and try boxcox
hist(log(yr2018$FgTa_Geomean))  # log clearly better

```

### 3a) DON
```{r}
## DON
don17_mod <- lm(DON ~ CO2, data = yr2017)
Anova(don17_mod, type = 3)
summary(don17_mod)

don18_mod <- lmer(sqrt(DON) ~ day * CO2 + (1|plot:CO2), data = yr2018)
Anova(don18_mod, type = 3)
# CO2 effect is not significant
summary(don18_mod)

```

### 3a) Fusarium biomass
```{r}
fg17_mod <- lm(FgTa_TRI6toPR1 ~ CO2, data = yr2017)
Anova(fg17_mod, type = 3)
summary(fg17_mod)

fg18_mod <- lmer(log(FgTa_Geomean) ~ day * CO2 + (1|plot:CO2), data = yr2018)
#fg18_mod.a <- lmer(log(FgTa_Geomean) ~ day * CO2 + (1|plot), data = yr2018)

Anova(fg18_mod, type = 3)
#Anova(fg18_mod.a, type = 3)

# CO2 effect is not significant
summary(fg18_mod)
#summary(fg18_mod.a)

```



```{r}
sessionInfo()
```



#### end